---
title: '[.data.table is very slow with a single column #5650'
output: html_document
date: "2024-02-02"
---
Regression
[.data.table is very slow with a single integer #5636
https://github.com/Rdatatable/data.table/issues/5636

fixed by new with optimization to allow avoid [ overhead #4488
https://github.com/Rdatatable/data.table/pull/4488


```{r}
library(data.table)
allIterations <- data.frame(v1 = runif(1e5), v2 = runif(1e5))
DoSomething <- function(row) someCalculation <- row[["v1"]] + 1
system.time(for (r in 1:nrow(allIterations)) DoSomething(allIterations[r, ]))
#   user  system elapsed 
#  3.384   0.007   3.392
allIterations <- as.data.table(allIterations)
setDTthreads(1)
system.time(for (r in 1:nrow(allIterations)) DoSomething(allIterations[r, , with=c(i=FALSE)]))
#   user  system elapsed 
#  5.432   0.121   5.554 
```


```{r}
atime.list <- atime::atime(
  N = nrow(newConfigurations),
  setup = {
    parameters <- list(types = c(p1 = "r", p2 = "r", p3 = "r", dummy = "c"),
                       digits = 4)
    n <- 10000
    newConfigurations <- data.table(p1 = runif(n), p2 = runif(n), p3 = runif(n),
                                    dummy = sample(c("d1", "d2"), n, replace = TRUE))
  },
)
```

```{r}
result.4488 <- atime::atime(
  N = 10^seq(1, 20),
  setup = {
    parameters <- list(types = c(p1 = "r", p2 = "r", p3 = "r", dummy = "c"),
                       digits = 4)
  },
  seconds.limit = 2,
  expr.list = list(
    newConfigurations = {
      data.table::data.table(p1 = runif(N), p2 = runif(N), p3 = runif(N),
                             dummy = sample(c("d1", "d2"), N, replace = TRUE))
    },
    dt.J = {
      for (i in seq_len(nrow(newConfigurations))) {
        set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))
      }
    },
    repair_sum2one_time = {
      for (i in seq_len(nrow(newConfigurations))) {
        repair_sum2one(as.data.frame(newConfigurations[i]), parameters)
      }
    }
  )
)
```



```{r}
result.4488 <- atime::atime(
  N=10^seq(1,20),
  setup = {
    parameters <- list(types = c(p1 = "r", p2 = "r", p3 = "r", dummy = "c"),
                       digits = 4)
    newConfigurations <- data.table(p1 = runif(N), p2 = runif(N), p3 = runif(N),
                                    dummy = sample(c("d1", "d2"), N, replace = TRUE))
  },
seconds.limit = 2,
expr.list = list(
  dt.J = for (i in seq_len(nrow(newConfigurations))) {
    set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))
  },
  repair_sum2one_time = 
    for (i in seq_len(nrow(newConfigurations))) {
      repair_sum2one(as.data.frame(newConfigurations[i]), parameters)
    }
  
  
)
)
```


```{r}
plot(result.4488)+
  labs(title = "data.table is very slow with a single column #5650 ")

png("result.4488.png")
plot(result.4488)+
  labs(title = "data.table is very slow with a single column #5650")
dev.off()
```


```{r}


library(atime)
library(ggplot2)
library(data.table)

```

```{r}
  tdir <- tempfile()
  dir.create(tdir)
  git2r::clone("https://github.com/Rdatatable/data.table", tdir)
  
```


```{r}

atime.list.4488 <- atime::atime_versions(
  pkg.path=tdir,
  pkg.edit.fun=function(old.Package, new.Package, sha, new.pkg.path){
    pkg_find_replace <- function(glob, FIND, REPLACE){
      atime::glob_find_replace(file.path(new.pkg.path, glob), FIND, REPLACE)
    }
    Package_regex <- gsub(".", "_?", old.Package, fixed=TRUE)
    Package_ <- gsub(".", "_", old.Package, fixed=TRUE)
    new.Package_ <- paste0(Package_, "_", sha)
    pkg_find_replace(
      "DESCRIPTION", 
      paste0("Package:\\s+", old.Package),
      paste("Package:", new.Package))
    pkg_find_replace(
      file.path("src","Makevars.*in"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      sprintf('packageVersion\\("%s"\\)', old.Package),
      sprintf('packageVersion\\("%s"\\)', new.Package))
    pkg_find_replace(
      file.path("src", "init.c"),
      paste0("R_init_", Package_regex),
      paste0("R_init_", gsub("[.]", "_", new.Package_)))
    pkg_find_replace(
      "NAMESPACE",
      sprintf('useDynLib\\("?%s"?', Package_regex),
      paste0('useDynLib(', new.Package_))
  },
  N=10^seq(1,20),
  setup={ 
    newConfigurations <- data.table(p1 = runif(N), p2 = runif(N), p3 = runif(N),
                                    dummy = sample(c("d1", "d2"), N, replace=TRUE))
    j <- colnames(newConfigurations)
for (i in seq_len(nrow(newConfigurations)))
      set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))
  },
  expr = data.table:::`[.data.table`(newConfigurations, for (i in seq_len(nrow(newConfigurations))) {
    set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))
  }),
  "master"="6f43a96ad478e02a48f61939e0c0d3047e949353",#https://github.com/Rdatatable/data.table/tree/6f43a96ad478e02a48f61939e0c0d3047e949353
  "PR4488"="c33e98bb0a2f919bc1b428a5b7985f45f9d88a77")#https://github.com/Rdatatable/data.table/pull/4488/commits/c33e98bb0a2f919bc1b428a5b7985f45f9d88a77
```
library(data.table)
parameters <- list(types = c(p1 = "r", p2 = "r", p3 = "r", dummy = "c"),
                   digits = 4)
n <- 10000
newConfigurations <- data.table(p1 = runif(n), p2 = runif(n), p3 = runif(n),
                                dummy = sample(c("d1", "d2"), n, replace=TRUE))

repair_sum2one <- function(configuration, parameters)
{
  isreal <- names(which(parameters$types[colnames(configuration)] == "r"))
  digits <- parameters$digits[isreal]
  c_real <- unlist(configuration[isreal])
  c_real <- c_real / sum(c_real)
  c_real[-1] <- round(c_real[-1], digits[-1])
  c_real[1] <- 1 - sum(c_real[-1])
  configuration[isreal] <- c_real
  return(configuration)
}
j <- colnames(newConfigurations)
for (i in seq_len(nrow(newConfigurations)))
      set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))

# Measure the time used by data.table operations
data.table_time <- system.time({
  j <- colnames(newConfigurations)
  for (i in seq_len(nrow(newConfigurations))) {
    set(newConfigurations, i, j = j, value = repair_sum2one(as.data.frame(newConfigurations[i]), parameters))
  }
})

# Measure the time used by the repair_sum2one function
repair_sum2one_time <- system.time({
  for (i in seq_len(nrow(newConfigurations))) {
    repair_sum2one(as.data.frame(newConfigurations[i]), parameters)
  }
})

# Print the execution times
print(data.table_time)
print(repair_sum2one_time)

```{r}
plot(atime.list.4488)+
  labs(title = "data.table is very slow with a single integer #5636")

png("atime.list.4488.png")
plot(atime.list.4488)+
  labs(title = "data.table is very slow with a single integer #5636")
dev.off()

```




