---
title: 'Selecting from data.table by row is very slow #3735'
output: html_document
date: "2024-02-02"
---


library(data.table)
allIterations <- data.frame(v1 = runif(1e5), v2 = runif(1e5))
DoSomething <- function(row) someCalculation <- row[["v1"]] + 1
system.time(for (r in 1:nrow(allIterations)) DoSomething(allIterations[r, ]))
#   user  system elapsed 
#  3.384   0.007   3.392
allIterations <- as.data.table(allIterations)
setDTthreads(1)
system.time(for (r in 1:nrow(allIterations)) DoSomething(allIterations[r, , with=c(i=FALSE)]))
#   user  system elapsed 
#  5.432   0.121   5.554 



```{r}
allIterations <- data.frame(v1 = runif(1e5), v2 = runif(1e5))
DoSomething <- function(row) {
  someCalculation <- row[["v1"]] + 1
}
system.time(
       {
         for (r in 1:nrow(allIterations)) {
           DoSomething(allIterations[r, ])
         }
       }
     )
##   user  system elapsed 
##   4.50    0.02    4.55 

allIterations <- as.data.table(allIterations)
system.time(
       {
         for (r in 1:nrow(allIterations)) {
           DoSomething(allIterations[r, ])
         }
       }
     )
##   user  system elapsed 
##   53.78   25.05   78.46 
```



```{r}


library(atime)
library(ggplot2)
library(data.table)

```

```{r}
  tdir <- tempfile()
  dir.create(tdir)
  git2r::clone("https://github.com/Rdatatable/data.table", tdir)
  
```


```{r}

atime.list.3735 <- atime::atime_versions(
  pkg.path=tdir,
  pkg.edit.fun=function(old.Package, new.Package, sha, new.pkg.path){
    pkg_find_replace <- function(glob, FIND, REPLACE){
      atime::glob_find_replace(file.path(new.pkg.path, glob), FIND, REPLACE)
    }
    Package_regex <- gsub(".", "_?", old.Package, fixed=TRUE)
    Package_ <- gsub(".", "_", old.Package, fixed=TRUE)
    new.Package_ <- paste0(Package_, "_", sha)
    pkg_find_replace(
      "DESCRIPTION", 
      paste0("Package:\\s+", old.Package),
      paste("Package:", new.Package))
    pkg_find_replace(
      file.path("src","Makevars.*in"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      sprintf('packageVersion\\("%s"\\)', old.Package),
      sprintf('packageVersion\\("%s"\\)', new.Package))
    pkg_find_replace(
      file.path("src", "init.c"),
      paste0("R_init_", Package_regex),
      paste0("R_init_", gsub("[.]", "_", new.Package_)))
    pkg_find_replace(
      "NAMESPACE",
      sprintf('useDynLib\\("?%s"?', Package_regex),
      paste0('useDynLib(', new.Package_))
  },
  N=10^seq(1,20),
  setup={ 
    allIterations <- data.frame(v1 = runif(N), v2 = runif(N))
    DoSomething <- function(row) {
      someCalculation <- row[["v1"]] + 1
      }
      allIterations <- as.data.table(allIterations)
      setDTthreads(1)
      
    },
  expr=data.table:::`[.data.table`({
    for (r in 1:nrow(allIterations)) {
      DoSomething(allIterations[r, ])
    }
  }),
  "master"="6f43a96ad478e02a48f61939e0c0d3047e949353",#https://github.com/Rdatatable/data.table/tree/6f43a96ad478e02a48f61939e0c0d3047e949353
  "PR4488"="45952d1619365a4624d939575a934583c33c8fbb")#https://github.com/Rdatatable/data.table/pull/4488/commits/c33e98bb0a2f919bc1b428a5b7985f45f9d88a77 
```


```{r}
plot(atime.list.3735)+
  labs(title = "Selecting from data.table by row is very slow #3735")

png("atime.list.3735.png")
plot(atime.list.3735)+
  labs(title = "Selecting from data.table by row is very slow #3735")
dev.off()

```
allIterations <- data.frame(v1 = runif(1e5), v2 = runif(1e5))
DoSomething <- function(row) {
  someCalculation <- row[["v1"]] + 1
}
system.time(
       {
         for (r in 1:nrow(allIterations)) {
           DoSomething(allIterations[r, ])
         }
       }
     )
##   user  system elapsed 
##   4.50    0.02    4.55 

allIterations <- as.data.table(allIterations)
system.time(
       {
         for (r in 1:nrow(allIterations)) {
           DoSomething(allIterations[r, ])
         }
       }
     )
##   user  system elapsed 
##   53.78   25.05   78.46 

```{r}
result.3735 <- atime::atime(
  
  N = 10^seq(1,7),
  setup = {
    allIterations <- data.frame(v1 = runif(N), v2 = runif(N))
    DoSomething <- function(row) {
      someCalculation <- row[["v1"]] + 1
      
    }
  },
  seconds.limit = 2,
  expr.list = list(
    DF = quote(atime({
      for (r in 1:nrow(allIterations)) {
        DoSomething(allIterations[r, ])
      }
    })),
    DT = quote(atime({
      for (r in 1:nrow(as.data.table(allIterations))) {
        DoSomething(as.data.table(allIterations)[r, ])
      }
    }))
  ),
  
)
```

```{r}
plot(result.3735)+
  labs(title = "Selecting from data.table by row is very slow #3735")

png("result.3735.png")
plot(result.3735)+
  labs(title = "Selecting from data.table by row is very slow #3735")
dev.off()
```
```{r}

atime.list.3735branches <- atime::atime_versions(
  pkg.path=tdir,
  pkg.edit.fun=function(old.Package, new.Package, sha, new.pkg.path){
    pkg_find_replace <- function(glob, FIND, REPLACE){
      atime::glob_find_replace(file.path(new.pkg.path, glob), FIND, REPLACE)
    }
    Package_regex <- gsub(".", "_?", old.Package, fixed=TRUE)
    Package_ <- gsub(".", "_", old.Package, fixed=TRUE)
    new.Package_ <- paste0(Package_, "_", sha)
    pkg_find_replace(
      "DESCRIPTION", 
      paste0("Package:\\s+", old.Package),
      paste("Package:", new.Package))
    pkg_find_replace(
      file.path("src","Makevars.*in"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      Package_regex,
      new.Package_)
    pkg_find_replace(
      file.path("R", "onLoad.R"),
      sprintf('packageVersion\\("%s"\\)', old.Package),
      sprintf('packageVersion\\("%s"\\)', new.Package))
    pkg_find_replace(
      file.path("src", "init.c"),
      paste0("R_init_", Package_regex),
      paste0("R_init_", gsub("[.]", "_", new.Package_)))
    pkg_find_replace(
      "NAMESPACE",
      sprintf('useDynLib\\("?%s"?', Package_regex),
      paste0('useDynLib(', new.Package_))
  },
  N=10^seq(1,20),
  setup={ 
    allIterations <- data.frame(v1 = runif(N), v2 = runif(N))
    DoSomething <- function(row) {
      someCalculation <- row[["v1"]] + 1
      }
      allIterations <- as.data.table(allIterations)
      setDTthreads(1)
      
    },
  expr=data.table:::`[.data.table`({
    for (r in 1:nrow(allIterations)) {
      DoSomething(allIterations[r, ])
    }
  }),
  "before"="d47a83fb2e25582e508f191f87a31ca81b736b57",#https://github.com/Rdatatable/data.table/commit/d47a83fb2e25582e508f191f87a31ca81b736b57
  "Regression"="469b9ebe16adabd5f7c68c2727f19c83fd444e43",#https://github.com/Rdatatable/data.table/pull/4488/commits/469b9ebe16adabd5f7c68c2727f19c83fd444e43
  "fixed"="45952d1619365a4624d939575a934583c33c8fbb") #https://github.com/Rdatatable/data.table/pull/4488/commits/45952d1619365a4624d939575a934583c33c8fbb
```



```{r}
plot(atime.list.3735branches)+
  labs(title = "Selecting from data.table by row is very slow #3735")

png("atime.list.3735branches.png")
plot(atime.list.3735branches)+
  labs(title = "Selecting from data.table by row is very slow #3735")
dev.off()
```

















